---
title: "Working with the Dryad Dataset Search Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with the Dryad Dataset Search Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview
This vignette demonstrates how to search for datasets on the Dryad repository using keywords, filter based on methods, and download dataset files for further analysis.

## Workflow

### 1. Load the Package
Make sure package is loaded.

```{r setup}
# Load your package
library(RC.SJ.package)
```

### 2. Search Dryad for Datasets
Use `search_datasets()` to query the Dryad API for datasets that contain specific keywords.

```{r}
results <- search_datasets(keyword = c("climate", "butterfly"), max_results = 50)

# Preview the results
head(results)
```

Each result includes the dataset's title, authors, publication year, DOI, abstract, and methods section (if available).

### 3. Filter Datasets by Method
Use `search_method()` to filter datasets that include specific keywords in the methods or abstract fields.

```{r}
filtered <-search_method(results, keywords = c("experiment, heatwave"), match_all = FALSE)

# View filtered results
filtered
```

If you would like results that contain all keywrods, use `match_all = TRUE`.

### 4. Download Dataset Files
Use `download_dataset()` to download the selected datasets (in .zip format) to your working directory.

```{r}
downloads <- download_dataset(filtered)

# View file paths to downloaded files
downloads
```

Each file is saved with a name related to its DOI. The function will issue a warning if a download fails.

### Notes
- All downloads will go to your current working directory.
- `max_results` caps how many results you retrieve from the Dryad API (deafult = 100)
- Some datasets do not include method sections, fitering will skip them
- This package uses base R and tidyverse packages like httr, jsonlite, dplyr, and stringr
